---
title: "Kaggle Titanic Competition"
author: "Mauro Vicic"
date: "3/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(mice)
library(DMwR)
library(randomForest)
```


First I loaded the data into two separate dataframes, usually I would merge them into one but when I was starting this project I was exploring the train dataset first so i just repeted the same steps to set up the test dataset.

```{r, include=FALSE}
titTrain = read_csv("KaggleComp/train.csv")
titTest = read_csv("KaggleComp/test.csv")
```

In order to fill in some unknown numeric values such as age and fare I used the knnImputation method.

```{r}
numsTrain <- sapply(titTrain, is.numeric)
numsTrain[2] = FALSE
newTitTrain = data.frame(titTrain[ ,numsTrain])
knnOutputTrain <- knnImputation(newTitTrain)

numsTest <- sapply(titTest, is.numeric)
newTitTest = data.frame(titTest[ ,numsTest])
knnOutputTest <- knnImputation(newTitTest)
```


After the knnImputation method was done I made sure that dataframes are clean and that there is no missing values in the character columns of interested. I did not use Name, Cabin and Ticket columns for this model.

```{r}
charTrain = sapply(titTrain,is.character)
charTrain[1] = TRUE
charTitTrain = data.frame(titTrain[ ,charTrain])

charTest = sapply(titTest,is.character)
charTest[1] = TRUE
charTitTest = data.frame(titTest[ ,charTest])

```


```{r}
fullTitanic = merge(knnOutputTrain, charTitTrain, by = "PassengerId")

fullTitanicTest = merge(knnOutputTest, charTitTest, by = "PassengerId")
```

```{r}
train = cbind(fullTitanic, titTrain[,2])
train[sapply(train, is.character)] <- lapply(train[sapply(train, is.character)], 
                                       as.factor)

fullTitanicTest[sapply(fullTitanicTest, is.character)] <- lapply(fullTitanicTest[sapply(fullTitanicTest, is.character)], 
                                       as.factor)
```

```{r}
summary(train$Embarked)
missing = which(is.na(train$Embarked))
train$Embarked[missing] = "S"
summary(train$Embarked)
```

```{r}
summary(fullTitanicTest$Embarked)
```


After the cleaning is done I made a random forest method which had a classification percentage of close to 79% which was good for top 31%.


```{r}
set.seed(7)
model <- randomForest(as.factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + Fare +
                                            Embarked,
                      data=train, 
                      importance=TRUE, 
                      ntree=500)
```

```{r}
varImpPlot(model)
```

```{r}
result = predict(model, fullTitanicTest)
```

```{r}
submit <- data.frame(PassengerId = fullTitanicTest$PassengerId, Survived = result)

write.csv(submit, file = "titanicSubmit.csv", row.names = FALSE)
```




